---
title: "Selecting color schemes for mapping ancestral states"
description:
  How to change the phytools default color scheme when visualizing the results of ancestral character state estimation
author:
  - first_name: "Joel"
    last_name: "Nitta"
    url: https://joelnitta.com
    affiliation: University of Tokyo
    affiliation_url: http://iwasakilab.k.u-tokyo.ac.jp/
    orcid_id: 0000-0003-4719-7472
date: 2021-06-02
preview: featured.png
creative_commons: CC BY
citation_url: https://joelnitta.com/color-scheme-anc-states 
repository_url: https://github.com/joelnitta/joelnitta-site
output:
  distill::distill_article:
    self_contained: false
params:
  slug: color-scheme-anc-states
  date: 2021-06-02
  repo: joelnitta/joelnitta-site
  site: https://joelnitta.com
editor_options: 
  chunk_output_type: console
---

<!--------------- setup post ----------------->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
refinery::renv_load(paste(params$date, params$slug, sep = "_"))
```

<!--------------- post ----------------->

The `phytools` package provides (among many other things) the [`contMap()`](https://rdrr.io/cran/phytools/man/contMap.html) function for estimating ancestral character states and [visualizing their changes along the branches of a phylogenetic tree](http://blog.phytools.org/search?q=contmap). It can either produce the plot directly (default), or be saved as an object with the `plot = FALSE` argument, to be further manipulated and plotted later with `plot()`.

## Default colors

I have to say I'm not a fan of the default color scheme, which is a rainbow palette going from red through yellow and green to blue. 

For example, let's [borrow some example code](http://www.phytools.org/eqg2015/asr.html) and look at the default plot:

```{r plot-contmap-default, layout = "l-body-outset"}
# code modified slightly from http://www.phytools.org/eqg2015/asr.html

## Load needed packages for this blogpost
library(phytools)
library(ggtree)
library(tidyverse)
library(scico)
library(viridisLite)

## Load anole tree
anole.tree <- read.tree("http://www.phytools.org/eqg2015/data/anole.tre")

## Load anole trait data, extract snout-vent-length (svl) as named vector
svl <- read_csv("http://www.phytools.org/eqg2015/data/svl.csv") %>%
  mutate(svl = set_names(svl, species)) %>%
  pull(svl)

# Plot with default color scheme
contmap_obj <- contMap(anole.tree, svl, plot = FALSE)

plot(
  contmap_obj, 
  type="fan", 
  legend = 0.7*max(nodeHeights(anole.tree)),
  fsize = c(0.5, 0.7))
```

Although this does provide a wide range of colors, **it's not obvious why one color is greater or less than the others**. In particular it's hard to discern the order of intermediate values (yellow, green, light blue). Indeed, there has been much written on [why the rainbow palette is generally not a good way to visualize continuous data](https://www.nature.com/articles/s41467-020-19160-7).

## Defining a new color palette

[`phytools::setMap()`](https://rdrr.io/cran/phytools/man/setMap.html) can be used to specify another color palette. `setMap()` passes its second argument (a vector of color names or [hexadecimals](https://en.wikipedia.org/wiki/Web_colors)) to [`colorRampPalette()`](https://rdrr.io/r/grDevices/colorRamp.html). `colorRampPalette()` is a bit unusual in that it's a function that produces a function, in this case, one that generates a vector of colors interpolating between the original input values:

```{r colorRampPalette}
# colorRampPalette() produces a function
my_color_func <- colorRampPalette(c("red", "yellow"))
class(my_color_func)

# The function generates n colors interpolating between
# the colors originally passed to colorRampPalette()
my_colors <- my_color_func(n = 6)
scales::show_col(my_colors)
```

So, this works fine for generating custom color gradients. But [designing accurate, color-blind friendly color palettes is not a simple task](https://www.fabiocrameri.ch/visualisation/). Fortunately, there are several packages available with such carefully crafted palettes. Two of my favorite are [`viridis`](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) and [`scico`](https://github.com/thomasp85/scico). **How can we use these with the plotting function in phytools?**

## Using `viridis` or `scico` palettes

Well, it turns out that as long as we specify the same number of colors, we can replicate the viridis color palette with `colorRampPalette()`. The only difference is the alpha, or transparency level, indicated at the [end of each hexidecimal with two letters](https://stackoverflow.com/questions/23201134/transparent-argb-hex-value) (here "FF"). There is no reason to use transparency here anyways, so that doesn't matter.

```{r colorRampPalette-viridis}
# viridis color palette with 6 colors
viridis(6)
# colorRampPalette() replicating viridis color palette
colorRampPalette(viridis(6))(6)
```

So here is the `viridis` version of the phytools plot:

```{r phytools-viridis, layout = "l-body-outset"}
# Count the number of unique character states in the observed data:
n_cols <- n_distinct(svl)

# Change the color palette
contmap_obj_viridis <- setMap(contmap_obj, viridis(n_cols))

# Plot the mapped characters with the new colors
plot(
  contmap_obj_viridis, 
  type="fan", 
  legend = 0.7*max(nodeHeights(anole.tree)),
  fsize = c(0.5, 0.7))
```

And here is another one, this time using a palette from `scico`:

```{r phytools-scico, layout = "l-body-outset"}
# Change the color palette
contmap_obj_scico <- setMap(contmap_obj, scico(n_cols, palette = "bilbao"))

# Plot the mapped characters with the new colors
plot(
  contmap_obj_scico, 
  type="fan", 
  legend = 0.7*max(nodeHeights(anole.tree)),
  fsize = c(0.5, 0.7))
```

I personally find this one even easier to interpret than `viridis`. It's very clear which values are low and high.

## `ggtree`

Just for completeness, here is code to replicate the plot in  [`ggtree`](https://github.com/YuLab-SMU/ggtree).

```{r plot-contmap-ggtree, layout = "l-body-outset"}
# Modified from https://yulab-smu.top/treedata-book/chapter4.html#color-tree

# Fit an ancestral state character reconstruction
fit <- phytools::fastAnc(anole.tree, svl, vars = TRUE, CI = TRUE)

# Make a dataframe with trait values at the tips
td <- data.frame(
  node = nodeid(anole.tree, names(svl)),
  trait = svl)

# Make a dataframe with estimated trait values at the nodes
nd <- data.frame(node = names(fit$ace), trait = fit$ace)

# Combine these with the tree data for plotting with ggtree
d <- rbind(td, nd)
d$node <- as.numeric(d$node)
tree <- full_join(anole.tree, d, by = 'node')

ggtree(
  tree, 
  aes(color = trait), 
  layout = 'circular', 
  ladderize = FALSE, continuous = "color", size = 1) +
  # >>> The important part! <<<
  # Choose your favorite scale_color_* function here: 
  scale_color_scico(palette = "bilbao") + 
  geom_tiplab(hjust = -.1, size = 2, color = "black") + 
  xlim(0, 1.2) + 
  theme(
    legend.position = c(0, .82),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8)
  ) 
```

That's it!

<!--------------- appendices ----------------->

```{r, echo=FALSE}
refinery::insert_appendix(
  repo_spec = params$repo, 
  name = paste(params$date, params$slug, sep = "_")
)
```
