---
title: "Publications"
date: "`r Sys.Date()`"
output: distill::distill_article
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(distilltools)
library(RefManageR)
library(bib2df)
library(janitor)
library(lubridate)
library(tidyverse)
```

```{r setup-refs}
# Load bibliography
bib_path <- "_bib/references.bib"
mybib <- ReadBib(bib_path, check = FALSE) 
mybib_df_raw <- bib2df(bib_path)

# Format dataframe version of bibliography
mybib_df <-
mybib_df_raw %>%
	clean_names() %>%
	remove_empty("cols") %>%
	select(key = bibtexkey, date, doi, title, author, journaltitle) %>%
  mutate(
  	date = parse_date_time(date, "ymd"),
  	year = year(date)) %>%
	arrange(desc(year), key)

# Set RefManageR options
oldopts <- BibOptions(
	style = "text",
	bib.style = "authoryear", 
	no.print.fields = c("doi", "url", "urldate"),
	dashed = FALSE, 
	sorting = "ydnt")
# Silently cite all references
purrr::walk(seq_along(mybib), ~NoCite(mybib, .x))
#
```

```{r cite-func}
doi_button <- function(key_select, bib_df) {
	doi <- filter(bib_df, key == key_select) %>% pull(doi)
	icon_link("doi", doi, glue("https://doi.org/{doi}"))
}
```

`r icon_link("google-scholar", "Google Scholar", "https://scholar.google.com/citations?hl=en&user=FtzrMYwAAAAJ")`  

## 2021

```{r, results = "asis"}
print(mybib[["Nitta2021"]])
```

`r doi_button("Nitta2021", mybib_df)`

## 2020

```{r, results = "asis"}
print(mybib[["Nitta2020"]])
```

`r doi_button("Nitta2020", mybib_df)`

```{r, results = "asis"}
print(mybib[["Nitta2020c"]])
```

`r doi_button("Nitta2020c", mybib_df)`

## 2019

```{r, results = "asis"}
print(mybib[["Ebihara2019a"]])
```

```{r, results = "asis"}
print(mybib[["Ebihara2019b"]])
```

`r doi_button("Ebihara2019b", mybib_df)`

```{r, results = "asis"}
print(mybib[["Nitta2019"]])
```

`r doi_button("Nitta2019", mybib_df)`

## 2018

```{r, results = "asis"}
print(mybib[["Gilbert2018"]])
```

`r doi_button("Gilbert2018", mybib_df)`

```{r, results = "asis"}
print(mybib[["Nitta2018"]])
```

`r doi_button("Nitta2018", mybib_df)`

## 2017

```{r, results = "asis"}
print(mybib[["Nitta2017"]])
```

`r doi_button("Nitta2017", mybib_df)`

```{r, results = "asis"}
print(mybib[["Pinson2016a"]])
```

`r doi_button("Pinson2016a", mybib_df)`

```{r, results = "asis"}
print(mybib[["Zhou2017"]])
```

`r doi_button("Zhou2017", mybib_df)`

## 2016

```{r, results = "asis"}
print(mybib[["Pouteau2016"]])
```

`r doi_button("Pouteau2016", mybib_df)`

## 2015

```{r, results = "asis"}
print(mybib[["Chen2015"]])
```

`r doi_button("Chen2015", mybib_df)`

## 2013

```{r, results = "asis"}
print(mybib[["Ebihara2013"]])
```

`r doi_button("Ebihara2013", mybib_df)`

## 2011

```{r, results = "asis"}
print(mybib[["Nitta2011"]])
```

`r doi_button("Nitta2011", mybib_df)`

```{r, results = "asis"}
print(mybib[["Nitta2011b"]])
```

`r doi_button("Nitta2011b", mybib_df)`

## 2010

```{r, results = "asis"}
print(mybib[["Ebihara2010"]])
```

`r doi_button("Ebihara2010", mybib_df)`

```{r, results = "asis"}
print(mybib[["Ebihara2010a"]])
```

## 2009

```{r, results = "asis"}
print(mybib[["Ebihara2009b"]])
```

`r doi_button("Ebihara2009b", mybib_df)`

```{r, results = "asis"}
print(mybib[["Nitta2009"]])
```

`r doi_button("Nitta2009", mybib_df)`

## 2008

```{r, results = "asis"}
print(mybib[["Nitta2008"]])
```

`r doi_button("Nitta2008", mybib_df)`

```{r, results = "asis"}
print(mybib[["Nitta2008a"]])
```

`r doi_button("Nitta2008a", mybib_df)`

